---
title: "Tidyverse data cleaning"
subtitle: "Coursera-course-project" 
title-block-banner: "#f0f3f5"
title-block-banner-color: "black"
author: "Ivan Osinnii"
date: "25 June 2024"
format: html
code-fold: true
editor: visual
toc: true
toc-location: left
number-sections: true
fig-cap-location: margin
css: style.css
---

# Load necessary libraries

```{r}
library(tidyverse) 
library(here) 
library(readr) 
library(DescTools) 
library(RColorBrewer) 
library(gridExtra)
```

# Read the data and pre-clean it

```{r}
getwd()
coverage <- read_csv('./../input/healthcare-coverage.csv', skip = 2) 
coverage <- na.omit(coverage)

spending <- read_csv('./../input/healthcare-spending.csv', skip =2) 
spending <- na.omit(spending)
```

# Read the data
```{r}

coverage <- read_csv('./../input/tidy_healthcare_coverage.csv')
spending <- read_csv('./../input/tidy_spending_clean_corrected.csv')
```

## View the structure of the datasets

Assuming both datasets have a common key for merging, e.g., 'State' or 'Year' Check for common columns

```{r}
glimpse(coverage)
glimpse(spending)
common_columns <- intersect(names(coverage), names(spending))
print(common_columns)
```

Merge the datasets on the common key

```{r}
merged_data <- left_join(coverage, spending, by = "Location") # Replace 'common_key' with the actual column name

```

## Subset columns for 2013 and 2014 year. Write intermediate datasets

```{r}

merged_data_2013 <- merged_data %>% 
  select(Location, Total_2013, Total_Health_Spending_2013)
write.csv(merged_data_2013, "./../input/merged_data_2013.csv",row.names = TRUE)

merged_data_2014 <- merged_data %>% 
  select(Location, Total_2014, Total_Health_Spending_2014)
write.csv(merged_data_2014, "./../input/merged_data_2014.csv")

```

# Analyze the relationship between healthcare 2013s coverage and healthcare spending

```{r}
plot <- ggplot(merged_data_2013, 
  aes(x = Total_2013, y = Total_Health_Spending_2013)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  # scale_color_manual(values = colorRampPalette(brewer.pal(9, "Reds"))(num_locations)) +  # Red palette accommodating all data points
  labs(title = "Total 2013 vs. Total Health Spending 2013",
       x = "Total Healthcare Coverage 2013",
       y = "Total Health Spending 2013")
```

## Analyze the relationship between healthcare 2014s coverage and healthcare spending

```{r}
plot2 <- ggplot(merged_data_2014, 
               aes(x = Total_2014, y = Total_Health_Spending_2014)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  # scale_color_manual(values = colorRampPalette(brewer.pal(9, "Reds"))(num_locations)) +  # Red palette accommodating all data points
  labs(title = "Total 2014 vs. Total Health Spending 2014",
       x = "Total Healthcare Coverage 2014",
       y = "Total Health Spending 2014")
```

```{r}
grid.arrange(plot, plot2, ncol = 2) # the correlation looks similar
```

#Conclusion It looks like the relationship between both parameters is linear in both 2013/2024, but there's more Total Health Spending in 2014 than in 2013
